---
title: "Adjacent Category Mixed Effects Modeling of Cuisine Ratings"
author: "Omar Lizardo & Clayton Childress"
format:
  html:
    toc: true
    embed-resources: true
    code-fold: true
    theme: cosmo
execute:
  message: false
  warning: false
---

## Introduction

This pipeline analyzes the ratings of different cuisines as a function of respondents' demographics, with a specific focus on their political ideological stance on social issues. We utilize a mixed-effects adjacent category framework to account for the bipolar 1-7 scale structure with a midpoint. 

The models group the data across all 15 evaluated cuisines, including random effects for both the individual `respondent` and the `cuisine` being evaluated. We fit and compare two models:
1. A strict (proportional odds) adjacent category model.
2. A generalized adjacent category model where the effect of political ideology is allowed to vary across adjacent thresholds (category-specific effects).

## Setup and Data Preparation

```{r setup}
#| message: false
#| warning: false
library(tidyverse)
library(haven)
library(brms)
library(tidybayes)
library(modelsummary)
library(here)

# Load data cleaning function
source(here("recode.dat.R"))
```

```{r data-prep}
# Load and preprocess the data
dat <- recode.dat()

# Create a unique respondent ID since we're going to pivot to long format
dat$respondent_id <- seq_len(nrow(dat))

# The 15 cuisines
cuisines_cols <- c("japanese", "french", "italian", "mexican", "moroccan", 
                   "korean", "peruvian", "native_american", "swedish", 
                   "pakistani", "ethiopian", "vietnamese", "nigerian", 
                   "jamaican", "lebanese")

# Select necessary variables
demographics <- c("age.f", "race.f", "gend.f", "educ.f", "inc.f", "city.f")

# Keep the continuous `social` variable as well as the categorical `spol.f`
# We will use `social` as the focal predictor for the category-specific effects.
dat_subset <- dat |>
  select(respondent_id, all_of(cuisines_cols), all_of(demographics), social, spol.f)

# Pivot to long format
dat_long <- dat_subset |>
  pivot_longer(
    cols = all_of(cuisines_cols),
    names_to = "cuisine",
    values_to = "rating"
  ) |>
  # Remove missing ratings
  filter(!is.na(rating), !is.na(social)) |>
  # Ensure rating is an ordered factor for brms ordinal models
  mutate(
    rating_ord = factor(rating, levels = 1:7, ordered = TRUE),
    cuisine = as.factor(cuisine),
    respondent_id = as.factor(respondent_id),
    # Mean-center the continuous social ideology variable for easier interpretation
    social_c = scale(social, center = TRUE, scale = FALSE)
  )

glimpse(dat_long)
```

## Exploratory Data Analysis

Let's look at the distribution of cuisine ratings across social ideology categories.

```{r eda-plot}
#| fig.width: 10
#| fig.height: 6
ggplot(dat_long, aes(x = rating_ord, fill = spol.f)) +
  geom_bar(position = "dodge") +
  facet_wrap(~cuisine, scales = "free_y") +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Cuisine Ratings (1-7) by Social Ideology",
    x = "Rating (1 = Traditional Home Recipe, 7 = Pro Chef at Restaurant)",
    y = "Count",
    fill = "Social Ideology"
  ) +
  theme_minimal()
```


## Mixed Effects Adjacent Category Models

We use the `brms` package to fit Bayesian ordinal mixed-effects models. The family `acat()` (adjacent category) is appropriate for responses that are discrete categories where we model the probabilities of adjacent categories. 

The baseline model enforces the "proportional odds" assumption (which in the context of `acat` means that the slope for our predictors is the same across all threshold boundaries).

### Model 1: Strict Adjacent Category Model

In this model, the effect of social ideology (`social_c`) and other demographics are constant across all adjacent rating thresholds.

```{r fit-model1}
#| eval: false
# Note: eval=FALSE to prevent long compilation times during initial render.
# Set eval=TRUE when ready to run the MCMC sampling.

formula_strict <- bf(
  rating_ord ~ 1 + social_c + age.f + race.f + gend.f + educ.f + inc.f + 
    (1 | respondent_id) + (1 | cuisine)
)

fit_strict <- brm(
  formula = formula_strict,
  data = dat_long,
  family = acat("logit"),
  prior = c(
    prior(normal(0, 1.5), class = "Intercept"),
    prior(normal(0, 1), class = "b"),
    prior(exponential(1), class = "sd")
  ),
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 1234,
  backend = "rstan"
)

# Save the model
saveRDS(fit_strict, file = here("fit_strict_acat.rds"))
```


### Model 2: Category-Specific Effects for Social Ideology

In this model, we relax the proportional odds assumption for social ideology by using `cs(social_c)`. This allows the effect of social ideology to vary across the different thresholds of the 1-7 scale (e.g., distinguishing between 1 vs 2 might have a different relationship with ideology than distinguishing between 6 vs 7).

```{r fit-model2}
#| eval: false

formula_cs <- bf(
  rating_ord ~ 1 + cs(social_c) + age.f + race.f + gend.f + educ.f + inc.f + 
    (1 | respondent_id) + (1 | cuisine)
)

fit_cs <- brm(
  formula = formula_cs,
  data = dat_long,
  family = acat("logit"),
  prior = c(
    prior(normal(0, 1.5), class = "Intercept"),
    prior(normal(0, 1), class = "b"),
    prior(exponential(1), class = "sd")
  ),
  chains = 4,
  cores = 4,
  iter = 2000,
  seed = 1234,
  backend = "rstan"
)

saveRDS(fit_cs, file = here("fit_cs_acat.rds"))
```


## Model Comparison

We compare the two models using LOO (Leave-One-Out cross-validation) to see if allowing the effect of social ideology to vary across thresholds improves model fit.

```{r compare-models}
#| eval: false
# Add loo criteria
fit_strict <- add_criterion(fit_strict, "loo")
fit_cs <- add_criterion(fit_cs, "loo")

# Compare
loo_comp <- loo_compare(fit_strict, fit_cs)
print(loo_comp)
```

## Extracting and Visualizing Effects

Once models are fit, we can plot the conditional effects, especially focusing on how social ideology interacts with the probability of choosing different ratings.

```{r visualize-effects}
#| eval: false
# Conditional effects for the non-proportional model
p_ce <- conditional_effects(fit_cs, effects = "social_c", categorical = TRUE)

plot(p_ce, plot = FALSE)[[1]] +
  labs(
    title = "Predicted Probabilities of Cuisine Ratings",
    subtitle = "Varying by Social Ideology",
    x = "Social Ideology (Mean Centered, Higher = More Conservative)",
    y = "Probability",
    color = "Rating"
  ) +
  theme_minimal()
```

