---
  title: "Cuisines MCA Analysis"
  author: "Omar Lizardo"
  date: "`r Sys.Date()`"
---

# Setup
```{r setup, include=FALSE}
    knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
    library(conflicted)
    library(here)
    library(dplyr)
    library(tidyr)
    library(ggplot2)
    library(sjPlot)
    library(haven)
    library(patchwork)
    library(factoextra)
    library(FactoMineR)
    library(ggrepel)
    library(stringr)
    library(corrplot)
    conflicts_prefer(dplyr::filter)
    conflicts_prefer(dplyr::select)
    conflicts_prefer(dplyr::summarize)
```

```{r Getting and Recoding Data}
    source(here("recode.dat.R"))
    dat <- recode.dat() |> 
        select(
            # dine_out (2 variables)
            fancy_restaurant, fast_food,
            # food_qual (6 variables)
            light, rich, authentic, familiar, big, small,
            # food (6 variables)
            caviar, oysters, nuggets, cheeseburger, sourdough, avocado,
            # authenticity (15 variables)
            japanese.f:lebanese.f,
            # quant.sup (7 variables)
            age, income, educ, peduc, arts, city, relig,
            # quali.sup (4 variables)
            race.f, sex.f, gend.f, pol.f
        ) |> 
        na.omit()
    names(dat)
```

```{r Multiple Factor Analysis}
    mfa.res <- MFA(dat, 
        group = c(2, 6, 6, 15, 7, 4), 
        type = c("c", "c", "c", "n", "c", "n"),
        name.group = c("dine_out", "food_qual", "food", "authenticity", "quant.sup", "quali.sup"), 
        graph = FALSE,
        num.group.sup = (5:6)
        )
    mfa.dat <- dat |> cbind(mfa.res$ind$coord)
```

```{r}
    cat.dat <- mfa.res$quali.var$coord |> 
        data.frame() |> 
        select(Dim.1, Dim.3)
    p <- ggplot(cat.dat, aes(x = Dim.1, y = Dim.3, label = rownames(cat.dat)))
    p <- p + geom_vline(xintercept = 0, color = "gray", linewidth = 1)
    p <- p + geom_hline(yintercept = 0, color = "gray", linewidth = 1)
    p <- p + geom_text_repel(
        size = 7, 
        max.overlaps = 30, 
        fontface = "bold", 
        alpha = 1
        )
    p + theme_minimal() + theme(text = element_text(size = 16))
```

```{r Eigenvalue Plot}
    d <- 10
    p <- fviz_eig(mfa.res, choice = "eigenvalue", ncp =d)
    p <- p + geom_hline(yintercept = 1, linetype = 2, color = "red", linewidth = 1)
    p <- p + theme(text = element_text(size = 16, face = "bold"))
    p + geom_text(label = paste0(round(mfa.res$eig[, 2], 1),"%")[1:d], vjust=-1.1, hjust = 0.3,size = 5) + ylim(0, 1.75)  + ggtitle("")
    save_plot(here("Plots", "eig.png"), height = 25, width = 20)
```

```{r}
    plot.mat <- get_mfa_var(mfa.res)$cor
    plot.mat_df <- plot.mat |> 
        as.data.frame() |> 
        tibble::rownames_to_column(var = "Variable") |> 
        pivot_longer(
            cols = -Variable,
            names_to = "Dimension",
            values_to = "Correlation"
        ) |> 
        mutate(Variable = factor(Variable, levels = rownames(plot.mat)))
    a <- 0.75
    b <- -0.62
    ggplot(plot.mat_df, aes(y = Variable, x = Dimension, fill = Correlation)) + 
        geom_tile(color = "white", lwd = 2, linetype = 1) + coord_fixed() +
        scale_fill_gradient2(
                    low = "darkred",
                    mid = "#FFFFCC",
                    high = "darkblue",
                    midpoint = 0, limits = c(b, a)) +
        geom_text(aes(label = sprintf("%.2f", Correlation)), 
                    vjust = 0.6, 
                    fontface = "bold", 
                    color = "white", size = 10) +
        scale_x_discrete(position = "top") +
        scale_y_discrete(limits = rev) +
        theme_minimal() +
        theme(
            axis.title = element_blank(),
            text = element_text(size = 22, face = "bold"),
            legend.position = "right",
            legend.title = element_blank(),
            legend.key.height = unit(3, "cm"),  
        ) 
    save_plot(here("Plots", "varcor.png"), height = 60, width = 40) 
```

```{r Regression Models}
    ind.vars <- c("Intercept", "Asian", "Black", "Hispanic", "Mixed Race (Other)", "Mixed Race (White)", "Woman", "Nonbinary/Other", "LGBTQ", "Education", "Parent's Educ.", "Child Arts Exp.", "Age", "Income", "City Size", "Conservative", "Liberal", "Religious Attendance")
    spec1 <- c("race.f", "gend.f" ,"sex.f", "educ", "peduc", "arts", "age", "income", "city", "pol.f", "relig")
    f1 <- as.formula(paste("Dim.1 ~", paste(spec1, collapse = " + ")))
    f2 <- as.formula(paste("Dim.2 ~", paste(spec1, collapse = " + ")))
    f3 <- as.formula(paste("Dim.3 ~", paste(spec1, collapse = " + ")))
    m1 <- lm(f1, data = mfa.dat)
    m2 <- lm(f2, data = mfa.dat)
    m3 <- lm(f3, data = mfa.dat)
    tab_model(m1, m2, m3,
            file = here("Tabs", "pca-reg-tab1.html"), 
            dv.labels = c("Fancy Taste <br> (Dim. 1)", 
                          "Common Taste <br> (Dim. 2)",
                          "Foodie Taste <br> (Dim. 3)"
                        ), 
            CSS = list(css.table = 'font-size: 12px;', css.td = 'border-bottom: 0px;'), 
            show.ci = FALSE,  show.se = TRUE, collapse.se = TRUE, p.style = "numeric_stars", show.std = "std2", show.est = FALSE, digits = 3,
            pred.labels = ind.vars,
            title = "Table 1. Coefficient estimates of ordinary least squares linear models predicting scores on the first three MFA dimensions."
            )
```

## Multivariate Analysis of Variance (MANOVA)
Here we test whether there are statistically significant differences in the multivariate centroids of the first three MFA dimensions across key demographic groups (race, gender, and political ideology) simultaneously.

```{r MANOVA}
    # Fit MANOVA model
    manova_model <- manova(
        cbind(Dim.1, Dim.2, Dim.3) ~ race.f + gend.f + pol.f, 
        data = mfa.dat
    )
    
    # Multivariate test results (Pillai's Trace is robust to assumption violations)
    print(summary(manova_model, test = "Pillai"))
    
    # Univariate follow-up ANOVAs
    print(summary.aov(manova_model))
```

## Linear Discriminant Analysis (LDA)
Having found that political ideology (`pol.f`) is a strong predictor of tastes in both regression and MANOVA, we use Linear Discriminant Analysis (LDA) to identify how the continuous taste dimensions (`Dim.1`, `Dim.2`, `Dim.3`) best discriminate between Liberals, Moderates, and Conservatives.

```{r LDA}
    library(MASS)
    
    # Fit the LDA model
    lda_model <- lda(
        pol.f ~ Dim.1 + Dim.2 + Dim.3, 
        data = mfa.dat
    )
    
    # Print the model results
    print(lda_model)
    
    # Predict discriminant scores
    lda_pred <- predict(lda_model)
    
    # Create an LDA score dataframe for plotting
    lda_plot_dat <- mfa.dat |> 
        mutate(
            LD1 = lda_pred$x[, 1],
            LD2 = lda_pred$x[, 2],
            predicted_pol = lda_pred$class
        )
    
    # Plot discriminant space
    ggplot(lda_plot_dat, aes(x = LD1, y = LD2, color = pol.f)) +
        geom_point(alpha = 0.4, size = 1.5) +
        stat_ellipse(linewidth = 1) +
        scale_color_manual(values = c("Liberal" = "blue", "Moderate" = "green", "Conservative" = "red")) +
        labs(
            title = "Linear Discriminant Analysis of Tastes by Political Ideology",
            x = "Linear Discriminant 1 (LD1)",
            y = "Linear Discriminant 2 (LD2)",
            color = "Ideology"
        ) +
        theme_minimal()
```

## Hierarchical Clustering on Principal Components (HCPC)
To discover distinct "taste profiles" or "food consumer subcultures" within the sample, we perform Hierarchical Clustering on Principal Components (HCPC) on our MFA individual coordinates.

```{r HCPC}
    # Run HCPC clustering (using 3 clusters)
    hcpc_res <- HCPC(mfa.res, nb.clust = 3, graph = FALSE)
    
    # Extract cluster membership back into our main dataset
    mfa.dat$Cluster <- hcpc_res$data.clust$clust
    
    # Visualize the hierarchical tree of clusters
    plot(hcpc_res, choice = "tree", main = "Hierarchical Clustering Tree")
    
    # Visualize the individuals on the principal component map colored by cluster
    plot(hcpc_res, choice = "map", main = "Individual Factor Map by Cluster")
```

### Profiling the Clusters using Multinomial Logit Analysis
To understand how socioeconomic, demographic, and ideological variables predict membership in these three taste clusters, we fit a Multinomial Logit Model predicting cluster membership.

To facilitate pattern detection and avoid visual clutter, we compute the predicted probabilities of being in each cluster across the life course (age) for different political ideologies, and present them in faceted plots for each of the clusters.

```{r Multinomial Logit}
    library(nnet)
    library(ggeffects)
    
    # Fit the Multinomial Logit Model
    multinom_model <- multinom(
        Cluster ~ pol.f + gend.f + educ + age, 
        data = mfa.dat, 
        trace = FALSE
    )
    
    # Print the model coefficients as Relative Risk Ratios (Exponentiated Coefficients)
    print(exp(coef(multinom_model)))
    
    # Calculate predicted probabilities across age and political ideology
    pred_res <- ggpredict(multinom_model, terms = c("age [all]", "pol.f"))
    pred_df <- as.data.frame(pred_res)
    
    # Plot predicted probabilities faceted by cluster to reduce visual clutter
    ggplot(pred_df, aes(x = x, y = predicted, color = group)) +
        geom_line(linewidth = 1.2) +
        facet_wrap(~response.level, labeller = labeller(response.level = function(x) paste("Cluster", x))) +
        scale_color_manual(values = c("Liberal" = "blue", "Moderate" = "green", "Conservative" = "red")) +
        labs(
            title = "Predicted Probabilities of Cluster Membership by Age and Political Ideology",
            x = "Age",
            y = "Predicted Probability",
            color = "Political Ideology"
        ) +
        theme_minimal() +
        theme(text = element_text(size = 14))
    
    # Save the faceted probability plot
    save_plot(here("Plots", "pred_prob_faceted.png"), height = 15, width = 25)
```