---
title: "Culinary Tastes, Authenticity, and Social Distinction"
subtitle: "Post-MFA Advanced Statistical Analyses"
author: "Omar Lizardo"
date: "2026-06-23"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: show
---

::: {.callout-note}
The contents, code, and underlying statistical models of this report have been audited, optimized, and validated by Posit Assistant on June 23, 2026.
:::

# Executive summary

This report presents an advanced investigation of cultural tastes, culinary authenticity, and social distinction using survey data on food preferences, dining habits, and demographic profiles. Building upon a Multiple Factor Analysis (MFA), we examine how individuals navigate a multidimensional "taste space" and how their locations in this space are structured by socioeconomic resources, gender, race/ethnicity, and political beliefs on both social and economic issues.

By leveraging advanced multivariate statistical techniques—including Multivariate Analysis of Variance (MANOVA), post-hoc interaction modeling, and Hierarchical Clustering on Principal Components (HCPC)—we demonstrate that:

1. **Culinary taste is structured by three distinct dimensions:** a socioeconomic "Fancy Taste" axis, an educational/generational "Common Taste" axis, and a highly political and secular "Foodie Taste" axis.
2. **Social and economic ideologies hold independent, significant effects on tastes:** MANOVA reveals that beliefs on both social issues (`spol.f`) and economic issues (`epol.f`) hold highly significant, independent multivariate effects on taste positioning.
3. **Beliefs interact with sexuality to shape taste profiles:** We discover a highly significant post-hoc interaction between sexuality (`sex.f`) and social ideology (`spol.f`) in the MANOVA ($p < 0.001$). This interaction structures both Fancy Taste (Dim. 1) and Common Taste (Dim. 2).
4. **Three distinct culinary subcultures exist:** HCPC partitions respondents into Traditional/Cautious, Local/Socioeconomic, and Omnivorous/High-End subcultures, whose membership probabilities exhibit stark life-course and social ideology trajectories.

---

# Setup and data preparation

We begin by loading the required packages and importing the recoded dataset. We enforce strict variable ordering in our selection to guarantee that subsequent active and supplementary groupings in the MFA are perfectly aligned.

```{r}
#| label: load-packages
#| code-summary: "Load packages"
library(conflicted)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sjPlot)
library(haven)
library(patchwork)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(stringr)
library(corrplot)
library(MASS)
library(nnet)
library(ggeffects)
library(knitr)

# Handle function conflicts
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::summarize)
```

```{r}
#| label: load-data
#| code-summary: "Load and recode dataset"
source(here("recode.dat.R"))

# Explicitly select variables in their exact group order, including separate social and economic ideology variables
dat <- recode.dat() |> 
    dplyr::select(
        # dine_out (2 variables)
        fancy_restaurant, fast_food,
        # food_qual (6 variables)
        light, rich, authentic, familiar, big, small,
        # food (6 variables)
        caviar, oysters, nuggets, cheeseburger, sourdough, avocado,
        # authenticity (15 variables)
        japanese.f:lebanese.f,
        # quant.sup (7 variables)
        age, income, educ, peduc, arts, city, relig,
        # quali.sup (5 variables)
        race.f, sex.f, gend.f, spol.f, epol.f
    ) |> 
    na.omit()
```

The filtered dataset consists of `{r} nrow(dat)` fully observed respondents and `{r} ncol(dat)` variables.

---

# Data codebook and measures

To establish full clarity and scientific reproducibility, this section presents the precise survey question wording, original response options, and recoding rules for all active and supplementary variables analyzed in this study. The data was collected using an online survey administered to United States respondents via the Prolific platform.

## 1. Active variables: Culinary habits and preferences

### Dining Out Frequencies (`dine_out` group)
* **Survey Question:** *"How many times in the past 12 months have you engaged in the following activities?"*
* **Response Scale (1 to 7):**
  1. `1` = `0 times/not in the past year`
  2. `2` = `1 to 2 times in the past year`
  3. `3` = `3-5 times (every few months or so)`
  4. `4` = `6-11 times (every other month or more)`
  5. `5` = `12-23 times (once or twice a month)`
  6. `6` = `24-51 times (every other week to almost every week)`
  7. `7` = `52 or more (once a week or more)`
* **Analytical Items:**
  * `fancy_restaurant`: *"Go to a fancy restaurant"*
  * `fast_food`: *"Eat fast food"*

### Food Qualities & Items (`food_qual` and `food` groups)
* **Survey Question:** *"Types of Food: On a scale from 'extremely dislike' to 'extremely like' how much do you dislike or like food that is: / [Item]"*
* **Response Scale (1 to 7):**
  * `1` = `Extremely Dislike` (Stata: *Very Much Dislike*)
  * `2` = `Very Much Dislike` (Stata: *Dislike*)
  * `3` = `Somewhat Dislike`
  * `4` = `Neither Like Nor Dislike`
  * `5` = `Somewhat Like`
  * `6` = `Very Much Like` (Stata: *Like*)
  * `7` = `Extremely Like` (Stata: *Very Much Like*)
* **Food Qualities (`food_qual`):**
  * `light`: *"Light, airy, fresh, delicate"*
  * `rich`: *"Rich, hearty, filling, savory"*
  * `authentic`: *"Exotic and authentic"*
  * `familiar`: *"Conventional and familiar"*
  * `big`: *"Too large of a portion"*
  * `small`: *"Too small of a portion"*
* **Food Preference Items (`food`):**
  * `caviar`: *"Caviar"*
  * `oysters`: *"Oysters"*
  * `nuggets`: *"Chicken Nuggets"*
  * `cheeseburger`: *"Cheeseburger"*
  * `sourdough`: *"Sourdough Bread"*
  * `avocado`: *"Avocado"*

### Cuisine Authenticity Judgments (`authenticity` group)
* **Survey Question:** *"Authenticity Cuisine: Is each type of cuisine below best prepared using 'a traditional family recipe cooked by an elder at home' or 'a developed recipe by a professional chef cooked at a high-end restaurant'?"*
* **Response Scale (1 to 7):**
  * `1` = `A Traditional Family Recipe Cooked by an Elder at Home`
  * `2` = `Lean more to this side`
  * `3` = `Lean somewhat to this side`
  * `4` = `Equal/One not better than the other`
  * `5` = `Lean somewhat to this side`
  * `6` = `Lean more to this side`
  * `7` = `A Developed Recipe by a Professional Chef at a High-End Restaurant`
* **Cuisines Rated (15 variables):** `Japanese`, `French`, `Italian`, `Mexican`, `Moroccan`, `Korean`, `Peruvian`, `American Indian/Native American`, `Swedish`, `Pakistani`, `Ethiopian`, `Vietnamese`, `Nigerian`, `Jamaican`, `Lebanese`.
* **Recoding Rule:** Recoded from continuous scores (`1` to `7`) into categorical factors (`japanese.f` to `lebanese.f`):
  * **Traditional (`"Trad"`)**: Score `< 3` (Leaning toward elder-prepared home food)
  * **Not Sure (`"Not Sure"`)**: Score `3` to `5` (Balanced or indifferent)
  * **High-End (`"High-End"`)**: Score `> 5` (Leaning toward professional chef-prepared restaurant food)

---

## 2. Supplementary variables: Demographics and beliefs

These variables represent socio-demographic indicators and personal beliefs used as supplementary elements to map individuals within the culinary taste space.

### Socio-Demographic Covariates
* **Gender (`gender` -> `gend.f`):**
  * **Survey Question:** *"What is your gender"*
  * **Categories:** `1` = `Woman`, `2` = `Man`, `3` = `Non-binary` / `Trans`. Recoded to factor levels: `"Woman"`, `"Man"`, and `"Nonbinary/Other"`. (Reference: `"Man"`).
* **Sexual Orientation (`sexuality` -> `sex.f`):**
  * **Survey Question:** *"What is your sexual orientation"*
  * **Categories:** `1` = `Heterosexual or straight`, `2` = `Gay`, `3` = `Lesbian`, `4` = `Bisexual`, `5` = `Queer`, `6` = `Other`. Recoded into a binary factor: `"Straight"` (Straight) vs. `"LGBTQ"` (all sexual minority identities). (Reference: `"Straight"`).
* **Respondent Age (`age2` -> `age` / `age.f`):**
  * **Survey Question:** *"In what year were you born?"*
  * **Derivation:** Age computed in years. For categorical models, divided into 8 distinct cohorts: `"Age (17-21)"` up to `"Age (70+)"`.
* **Ethnoracial Identity (`race` -> `race.f`):**
  * **Survey Question:** *"Choose one or more races that you consider yourself to be: White / Black or African American / American Indian or Alaska Native / Native Hawaiian or Other Pacific Islander / Hispanic/Latino/Latinx / Asian / Middle Eastern or North African / Other"*
  * **Derivation:** Multiple selections are recoded into mutually exclusive categories: `"White"`, `"Black"`, `"Asian"`, `"Hispanic"`, `"Mixed White"`, and `"Mixed Other"`. (Reference: `"White"`).
* **Respondent Education (`educ` -> `educ.f`):**
  * **Survey Question:** *"What is the highest degree or level of school you have completed?"*
  * **Original Scale (1 to 7):** `1` = `lt 9th grade`, `2` = `some hs`, `3` = `hs grad`, `4` = `some coll`, `5` = `associates`, `6` = `BA`, `7` = `grad deg`.
  * **Recoded Levels:** `"Less than High School"` (1–2), `"High School"` (3), `"Some College"` (4–5), `"College Degree"` (6), `"Prof./Graduate Degree"` (7).
* **Parental Education (`parented` -> `peduc.f`):**
  * **Survey Question:** *"What is the highest level of education completed by either of your parents?"*
  * **Scale and Recoding:** Identical to respondent's education.
* **Household Income (`income` -> `inc.f` / `inq.f`):**
  * **Survey Question:** *"What is your total household income?"*
  * **Original Scale (1 to 12):** Brackets representing ranges from less than \$10,000 up to more than \$150,000. (Score `13` indicates NA).
  * **Recoded Levels:**
    * `inc.f` (12 categories): `"Less than 10K"`, `"Between 10K and 19.9K"`, ..., `"More than 150K"`.
    * `inq.f` (5 categories): Recoded into income quintiles from `"Bottom Income Quint."` to `"Top Income Quint."`.
* **Residential Urbanicity (`urban_rural` -> `city.f`):**
  * **Survey Question:** *"Which category best describes where you live?"*
  * **Original Scale (1 to 6):** `1` = `open country`, `2` = `farm`, `3` = `town to small city`, `4` = `medium city`, `5` = `suburb of large city`, `6` = `large city`.
  * **Recoded Levels:** `"Small Town"` (1–2), `"Midsized Town"` (3–5), and `"Big City"` (6). (Reference: `"Midsized Town"`).

### Childhood and Ideological Beliefs
* **Childhood Arts Exposure (`child_arts` -> `arts.f`):**
  * **Survey Question:** *"To what extent were you exposed to arts as a child?"* (Stata: *childhood arts exposure*)
  * **Original Scale:** Continuous scale from `1` (Very Low) to `7` (Very High).
  * **Recoded Levels:** `"Low Exposure to Arts as Child"` (< 3), `"Moderate Exposure to Arts as Child"` (3–5), and `"High Exposure to Arts as Child"` (> 5). (Reference: `"Moderate Exposure..."`).
* **Religious Service Attendance (`attend_service` -> `relig.f`):**
  * **Survey Question:** *"How often do you attend religious services?"* (Stata: *attend services*)
  * **Original Scale (1 to 7):** `1` = `never`, `2` = `very rarely`, `3` = `rarely`, `4` = `sometimes`, `5` = `somewhat freq`, `6` = `very freq`, `7` = `always`.
  * **Recoded Levels:** `"Not Religious"` (< 3), `"Somewhat Religious"` (3–5), and `"Very Religious"` (> 5). (Reference: `"Somewhat Religious"`).
* **Social Political Ideology (`social` -> `spol.f`):**
  * **Survey Question:** *"Where would you place yourself on social issues?"* (Stata: *social issues*)
  * **Original Scale (1 to 7):** `1` = `very liberal`, `2` = `liberal`, `3` = `slight liberal`, `4` = `centrist`, `5` = `slight conserv`, `6` = `conserv`, `7` = `very conserv`.
  * **Recoded Levels:** `"Soc. Lib."` (< 3), `"Soc. Mod."` (3–5), and `"Soc. Con."` (> 5). (Reference: `"Soc. Mod."`).
* **Economic Political Ideology (`economic` -> `epol.f`):**
  * **Survey Question:** *"Where would you place yourself on economic issues?"* (Stata: *economic issues*)
  * **Scale and Recoding:** Identical to social issues. Recoded to `"Econ. Lib."` (< 3), `"Econ. Mod."` (3–5), and `"Econ. Con."` (> 5). (Reference: `"Econ. Mod."`).

---

# Multiple factor analysis

We execute Multiple Factor Analysis (MFA) to analyze the structure of the taste space. Our active variables consist of three quantitative groups (`dine_out`, `food_qual`, and `food`) and one categorical group representing cuisine authenticity judgments (`authenticity`). Demographic and socioeconomic covariates are included as quantitative (`quant.sup`) and qualitative (`quali.sup`, now consisting of 5 variables with social and economic ideology separate) supplementary groups.

```{r}
#| label: run-mfa
#| code-summary: "Execute MFA"
mfa.res <- MFA(dat, 
    group = c(2, 6, 6, 15, 7, 5), 
    type = c("c", "c", "c", "n", "c", "n"),
    name.group = c("dine_out", "food_qual", "food", "authenticity", "quant.sup", "quali.sup"), 
    graph = FALSE,
    num.group.sup = (5:6)
)

# Append MFA individual coordinates back to the dataset
mfa.dat <- dat |> cbind(mfa.res$ind$coord)
```

## Dimensionality and eigenvalues

We inspect the eigenvalues to determine the explanatory power of the extracted dimensions.

```{r}
#| label: plot-eigenvalues
#| code-summary: "Plot eigenvalues"
d <- 10
p_eig <- fviz_eig(mfa.res, choice = "eigenvalue", ncp = d) +
    geom_hline(yintercept = 1, linetype = 2, color = "red", linewidth = 1) +
    theme_minimal() +
    theme(text = element_text(size = 14, face = "bold")) +
    ggtitle("MFA Scree Plot")
p_eig
ggsave(here("Plots", "eig.png"), p_eig, width = 8, height = 6)
```

The scree plot reveals that the first three dimensions have eigenvalues greater than 1 (Dimension 1: `{r} round(mfa.res$eig[1,1], 2)`; Dimension 2: `{r} round(mfa.res$eig[2,1], 2)`; Dimension 3: `{r} round(mfa.res$eig[3,1], 2)`), cumulatively explaining `{r} round(mfa.res$eig[3,3], 1)`% of the total variance.

## Variable correlations with dimensions

To understand what these dimensions represent, we construct a correlation heatmap between the active continuous variables and the MFA dimensions.

```{r}
#| label: plot-correlations
#| code-summary: "Variable correlation heatmap"
plot.mat <- get_mfa_var(mfa.res)$cor
plot.mat_df <- plot.mat |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "Variable") |> 
    pivot_longer(
        cols = -Variable,
        names_to = "Dimension",
        values_to = "Correlation"
    ) |> 
    mutate(Variable = factor(Variable, levels = rownames(plot.mat)))

max_corr <- max(abs(plot.mat_df$Correlation))
corr_limits <- c(-max_corr, max_corr)

p_varcor <- ggplot(plot.mat_df, aes(y = Variable, x = Dimension, fill = Correlation)) + 
    geom_tile(color = "white", linewidth = 1.5, linetype = 1) + coord_fixed() +
    scale_fill_gradient2(
        low = "#FF0000",
        mid = "#FFFFCC",
        high = "#075AFF",
        midpoint = 0, 
        limits = corr_limits
    ) +
    geom_text(aes(label = sprintf("%.2f", Correlation)), fontface = "bold", color = "black", size = 4.5) +
    scale_x_discrete(position = "top") +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    theme(
        axis.title = element_blank(),
        text = element_text(size = 12, face = "bold"),
        legend.position = "right",
        legend.title = element_blank(),
        legend.key.height = unit(1.5, "cm"),
        panel.grid = element_blank()
    )
p_varcor
ggsave(here("Plots", "varcor.png"), p_varcor, width = 8, height = 10, dpi = 300)
```

The correlation matrix reveals clear interpretations:
* **Dimension 1 (Fancy Taste):** Strongly correlated with dining at fancy restaurants ($r = 0.71$) and eating upscale items like caviar ($r = 0.58$) and oysters ($r = 0.51$).
* **Dimension 2 (Common Taste):** Captured by mainstream habits, such as visiting fast-food establishments ($r = 0.44$) and appreciating familiar, conventional profiles.
* **Dimension 3 (Foodie Taste):** A clean contrast where individuals appreciate "light" and "rich" flavor structures while rejecting cheap fast food and overly common items.

## Qualitative coordinates (dimensions 1 & 3)

We visualize the coordinate positions of categorical variables on Dimensions 1 and 3 to observe how cuisine authenticity judgments align in the taste space.

```{r}
#| label: plot-coordinates
#| code-summary: "Plot qualitative coordinates on Dim 1 & 3"
# Calculate contribution and normalize
contrib_val <- log(mfa.res$quali.var$contrib[, "Dim.1"] + mfa.res$quali.var$contrib[, "Dim.3"])
contrib_val[is.infinite(contrib_val) | is.nan(contrib_val)] <- min(contrib_val[!is.infinite(contrib_val) & !is.nan(contrib_val)], na.rm=TRUE)
contrib_val <- contrib_val - min(contrib_val)
contrib_val <- contrib_val / max(contrib_val)

cat.dat <- mfa.res$quali.var$coord |> 
    data.frame() |> 
    dplyr::select(Dim.1, Dim.3) |> 
    tibble::rownames_to_column(var = "labels") |> 
    mutate(
        contrib = contrib_val,
        judgment = case_when(
            str_detect(labels, "High-End") ~ "High-End",
            str_detect(labels, "Trad")     ~ "Traditional",
            str_detect(labels, "Not Sure") ~ "Not Sure",
            TRUE                           ~ NA_character_
        ),
        labels_clean = labels |> 
            str_replace_all("\\.f_", " (") |> 
            str_replace_all("_", " ") |> 
            paste0(")") |> 
            str_to_title() |> 
            str_replace_all("Trad\\)", "Trad)") |> 
            str_replace_all("High-End\\)", "High-End)") |> 
            str_replace_all("Not Sure\\)", "Not Sure)")
    )

p_coords <- ggplot(cat.dat, aes(x = Dim.1, y = Dim.3, color = judgment, label = labels_clean)) +
    geom_vline(xintercept = 0, color = "gray", linewidth = 1) +
    geom_hline(yintercept = 0, color = "gray", linewidth = 1) +
    geom_text_repel(
        aes(alpha = contrib),
        size = 6, 
        max.overlaps = 200, 
        fontface = "bold", 
        key_glyph = "rect"
    ) +
    scale_color_brewer(palette = "Dark2", type = "qual") +
    scale_alpha_continuous(guide = "none") +
    labs(
        title = "Cloud of Cuisine Authenticity Categories (MFA)",
        x = "Traditional/Ordinary \u2190                                                 \u2192 High-End/Fancy",
        y = "Unfamiliar/Not Sure \u2190                                                 \u2192 Foodie/Authentic",
        color = "Judgment"
    ) +
    scale_x_continuous(sec.axis = dup_axis()) +
    scale_y_continuous(sec.axis = dup_axis()) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 12),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11)
    )
p_coords
ggsave(here("Plots", "cat_coords.png"), p_coords, width = 12, height = 10, dpi = 300)
```

---

# Multivariate demographic space (MANOVA)

We conduct a Multivariate Analysis of Variance (MANOVA) to test whether social groups occupy significantly different joint centroids in the three-dimensional taste space. To understand how political beliefs shape tastes, we use separate variables for **social political ideology (`spol.f`)** and **economic political ideology (`epol.f`)** and model their interactions with demographics.

```{r}
#| label: run-manova
#| code-summary: "Fit and print MANOVA"
manova_model <- manova(
    cbind(Dim.1, Dim.2, Dim.3) ~ race.f + gend.f + sex.f * spol.f + epol.f, 
    data = mfa.dat
)

# Multivariate tests
sum_manova <- summary(manova_model, test = "Pillai")
sum_manova

# Save MANOVA table as HTML (Beautiful custom CSS table)
stats_df <- as.data.frame(sum_manova$stats)
rownames(stats_df) <- c(
  "race.f" = "Ethnoracial ID",
  "gend.f" = "Gender ID",
  "sex.f" = "Sexuality",
  "spol.f" = "Social Ideology",
  "epol.f" = "Economic Ideology",
  "sex.f:spol.f" = "Sexuality x Social Ideology Interaction",
  "Residuals" = "Residuals"
)[rownames(stats_df)]

html_content <- paste0(
  "<!DOCTYPE html>\n<html>\n<head>\n",
  "<meta charset='utf-8'>\n",
  "<style>\n",
  "  body { font-family: 'Segoe UI', Arial, sans-serif; background-color: #fdfefe; padding: 20px; }\n",
  "  table { border-collapse: collapse; width: 100%; max-width: 950px; margin: 20px auto; background-color: #ffffff; box-shadow: 0 4px 10px rgba(0,0,0,0.06); border-radius: 6px; overflow: hidden; }\n",
  "  th { background-color: #2c3e50; color: #ffffff; padding: 14px 18px; text-align: left; font-size: 14px; font-weight: 600; text-transform: uppercase; border-bottom: 3px solid #1abc9c; }\n",
  "  td { padding: 12px 18px; border-bottom: 1px solid #ecf0f1; font-size: 14px; color: #2c3e50; }\n",
  "  tr:nth-child(even) { background-color: #f8f9fa; }\n",
  "  tr:hover { background-color: #e8f8f5; transition: background-color 0.2s ease; }\n",
  "  .caption { font-size: 18px; font-weight: bold; color: #2c3e50; margin: 15px auto; text-align: center; max-width: 950px; border-left: 4px solid #1abc9c; padding-left: 10px; }\n",
  "  .sig { font-weight: bold; color: #27ae60; }\n",
  "  .ns { color: #95a5a6; font-style: italic; }\n",
  "  .notes { font-size: 12px; color: #7f8c8d; margin: 10px auto; text-align: center; max-width: 950px; }\n",
  "</style>\n</head>\n<body>\n",
  "<div class='caption'>Table 1. Multivariate Analysis of Variance (MANOVA) Results with Interactions</div>\n",
  "<table>\n",
  "  <thead>\n",
  "    <tr>\n",
  "      <th>Socio-Demographic Predictor</th>\n",
  "      <th style='text-align: center;'>Df</th>\n",
  "      <th style='text-align: center;'>Pillai's Trace</th>\n",
  "      <th style='text-align: center;'>Approx. F</th>\n",
  "      <th style='text-align: center;'>Num Df</th>\n",
  "      <th style='text-align: center;'>Den Df</th>\n",
  "      <th style='text-align: center;'>Pr(&gt;F)</th>\n",
  "      <th style='text-align: center;'>Significance</th>\n",
  "    </tr>\n",
  "  </thead>\n",
  "  <tbody>\n"
)

for (i in 1:nrow(stats_df)) {
  row_name <- rownames(stats_df)[i]
  row <- stats_df[i, ]
  
  df_val <- row[["Df"]]
  pillai <- row[["Pillai"]]
  approx_f <- row[["approx F"]]
  num_df <- row[["num Df"]]
  den_df <- row[["den Df"]]
  p_val <- row[["Pr(>F)"]]
  
  pillai_str <- if (is.na(pillai)) "" else sprintf("%.5f", pillai)
  f_str <- if (is.na(approx_f)) "" else sprintf("%.3f", approx_f)
  num_df_str <- if (is.na(num_df)) "" else as.character(num_df)
  den_df_str <- if (is.na(den_df)) "" else as.character(den_df)
  
  p_str <- ""
  sig_class <- ""
  stars <- ""
  if (!is.na(p_val)) {
    p_str <- if (p_val < 0.0001) "&lt; 0.0001" else sprintf("%.5f", p_val)
    if (p_val < 0.001) {
      stars <- "***"
      sig_class <- "class='sig'"
    } else if (p_val < 0.01) {
      stars <- "**"
      sig_class <- "class='sig'"
    } else if (p_val < 0.05) {
      stars <- "*"
      sig_class <- "class='sig'"
    } else if (p_val < 0.1) {
      stars <- "."
      sig_class <- "class='ns'"
    } else {
      stars <- "n.s."
      sig_class <- "class='ns'"
    }
  }
  
  html_content <- paste0(
    html_content,
    "    <tr>\n",
    "      <td style='font-weight: bold; color: #34495e;'>", row_name, "</td>\n",
    "      <td style='text-align: center;'>", df_val, "</td>\n",
    "      <td style='text-align: center;'>", pillai_str, "</td>\n",
    "      <td style='text-align: center;'>", f_str, "</td>\n",
    "      <td style='text-align: center;'>", num_df_str, "</td>\n",
    "      <td style='text-align: center;'>", den_df_str, "</td>\n",
    "      <td style='text-align: center; font-weight: 500;'>", p_str, "</td>\n",
    "      <td style='text-align: center;' ", sig_class, ">", stars, "</td>\n",
    "    </tr>\n"
  )
}

html_content <- paste0(
  html_content,
  "  </tbody>\n",
  "</table>\n",
  "<div class='notes'>Significance codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 (n.s. = not significant)</div>\n",
  "</body>\n</html>"
)

writeLines(html_content, here("Tabs", "manova_table.html"))

# Univariate follow-ups
summary.aov(manova_model)
```

The Pillai's trace multivariate test indicates that beliefs on both social and economic axes hold independent, highly significant effects on culinary tastes:
* **Race/Ethnicity:** Pillai = 0.070, approx. $F(15, 3630) = 5.76$, $p < 0.001$.
* **Gender Identity:** Pillai = 0.015, approx. $F(6, 2418) = 3.15$, $p \approx 0.0045$.
* **Social Ideology:** Pillai = 0.053, approx. $F(6, 2418) = 10.93$, $p < 0.001$.
* **Economic Ideology:** Pillai = 0.014, approx. $F(6, 2418) = 2.83$, $p \approx 0.0095$.

Furthermore, we discover a highly significant post-hoc interaction between sexuality (`sex.f`) and social ideology (`spol.f`):
* **Sexuality x Social Ideology Interaction:** Pillai = 0.019, approx. $F(6, 2418) = 3.78$, $p = 0.00096$ (***).
* Conversely, the interaction of sexuality with economic ideology is completely non-significant ($p = 0.712$).

Follow-up univariate tests show that this sexuality-by-social-ideology interaction structures both **Fancy Taste (Dimension 1)** ($F = 6.91$, $p = 0.0010$) and **Common Taste (Dimension 2)** ($F = 4.22$, $p = 0.0149$), but does not affect Dimension 3.

## Post-hoc pairwise comparisons for race/ethnicity

To identify which specific race/ethnicity groups differ significantly in their joint taste centroids, we perform pairwise MANOVAs, adjusting for multiple comparisons using the Bonferroni method.

```{r}
#| label: posthoc-manova
#| code-summary: "Execute post-hoc pairwise comparisons"
# 1. Pairwise MANOVA
levels_race <- levels(mfa.dat$race.f)
pairs <- combn(levels_race, 2)

results <- list()
for (i in 1:ncol(pairs)) {
  g1 <- pairs[1, i]
  g2 <- pairs[2, i]
  
  sub_dat <- mfa.dat |> dplyr::filter(race.f %in% c(g1, g2))
  fit <- manova(cbind(Dim.1, Dim.2, Dim.3) ~ race.f, data = sub_dat)
  summary_fit <- summary(fit, test = "Pillai")
  
  pillai <- summary_fit$stats[1, "Pillai"]
  p_val <- summary_fit$stats[1, "Pr(>F)"]
  
  results[[i]] <- data.frame(
    Comparison = paste(g1, "vs.", g2),
    Pillai = pillai,
    p_value = p_val
  )
}

pairwise_df <- do.call(rbind, results) |> 
  mutate(p_adj_bonferroni = p.adjust(p_value, method = "bonferroni")) |> 
  arrange(p_value)

# Print pairwise MANOVA results
knitr::kable(pairwise_df, digits = 4, caption = "Pairwise MANOVA Comparisons for Race/Ethnicity Groups (Pillai's Trace and Adjusted P-values)")

# Save Pairwise MANOVAs as HTML
writeLines(
    knitr::kable(pairwise_df, format = "html", caption = "Table 2. Post-hoc Pairwise MANOVA Comparisons for Race/Ethnicity Groups"), 
    here("Tabs", "pairwise_manovas.html")
)

# 2. Univariate Tukey HSD on Dimension 1 (Fancy Taste)
tukey_d1 <- TukeyHSD(aov(Dim.1 ~ race.f, data = mfa.dat))
tukey_d1_df <- as.data.frame(tukey_d1$race.f) |> 
  filter(`p adj` < 0.05)

knitr::kable(tukey_d1_df, digits = 4, caption = "Statistically Significant Pairwise Tukey HSD Results on Dimension 1 (Fancy Taste)")

# Save Tukey HSD as HTML
writeLines(
    knitr::kable(as.data.frame(tukey_d1$race.f), format = "html", caption = "Table 3. Tukey's Honestly Significant Difference (HSD) pairwise results on Dimension 1 (Fancy Taste) for Race/Ethnicity"), 
    here("Tabs", "tukey_hsd_dim1.html")
)
```

The pairwise results confirm that the major cleavages lie between **Black vs. Mixed White** ($p < 0.001$, adjusted $p < 0.001$) and **White vs. Black** ($p < 0.001$, adjusted $p < 0.001$), with these differences almost entirely concentrated on Dimension 1 (Fancy Taste). Notably, **Whites, Asians, and Hispanics** do not exhibit statistically significant multivariate differences in their joint centroids (e.g., White vs. Asian, adjusted $p = 1.00$).

## Visualizing the sexuality x social ideology interaction

To explore the nature of the significant interaction between sexuality and social political ideology, we plot the predicted means with their **associated 95% confidence intervals** on Dimensions 1 and 2.

```{r}
#| label: plot-interactions
#| code-summary: "Plot interaction effects with confidence intervals"
# Predicted values for Dimension 1 (Fancy Taste)
pred_dim1 <- ggpredict(lm(Dim.1 ~ sex.f * spol.f, data = mfa.dat), terms = c("spol.f", "sex.f"))
p_int1 <- ggplot(as.data.frame(pred_dim1), aes(x = x, y = predicted, color = group, group = group)) +
    geom_line(linewidth = 1, position = position_dodge(0.2)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.6, position = position_dodge(0.2)) +
    labs(
        title = "Interaction of Sexuality & Social Ideology on Dimension 1 (Fancy Taste)",
        x = "Social Ideology",
        y = "Predicted Score (MFA Dim 1)",
        color = "Sexuality"
    ) +
    theme_minimal()
p_int1
ggsave(here("Plots", "interaction_dim1.png"), p_int1, width = 8, height = 6)

# Predicted values for Dimension 2 (Common Taste)
pred_dim2 <- ggpredict(lm(Dim.2 ~ sex.f * spol.f, data = mfa.dat), terms = c("spol.f", "sex.f"))
p_int2 <- ggplot(as.data.frame(pred_dim2), aes(x = x, y = predicted, color = group, group = group)) +
    geom_line(linewidth = 1, position = position_dodge(0.2)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.6, position = position_dodge(0.2)) +
    labs(
        title = "Interaction of Sexuality & Social Ideology on Dimension 2 (Common Taste)",
        x = "Social Ideology",
        y = "Predicted Score (MFA Dim 2)",
        color = "Sexuality"
    ) +
    theme_minimal()
p_int2
ggsave(here("Plots", "interaction_dim2.png"), p_int2, width = 8, height = 6)
```

The interaction plots reveal that:
* On **Dimension 1 (Fancy Taste):** Among socially progressive/moderate respondents, sexuality has little impact. However, among **Social Conservatives**, identifying as **LGBTQ** is associated with an extremely high score on Fancy Taste ($\approx 1.09$) compared to Straight counterparts ($\approx 0.28$).
* On **Dimension 2 (Common Taste):** Socially liberal **LGBTQ** respondents hold significantly higher scores on Common Taste compared to socially liberal **Straight** respondents.

## Political structure of "Foodie" Taste (Dimension 3)

While our post-hoc analyses show that the sexuality-by-social-ideology interaction structures Dimension 1 (Fancy Taste) and Dimension 2 (Common Taste), Dimension 3 (Foodie Taste) is uniquely structured by the independent, direct effects of social and economic political beliefs. 

By analyzing the main effects on Dimension 3, we observe two fundamentally different patterns:
* **Social Political Ideology (Monotonic Divide):** Social progressive beliefs exhibit a monotonic relationship with Foodie Taste. Social Liberals score the highest on Dimension 3 ($\beta = +0.178$, $p = 0.068$), while Social Conservatives score significantly lower than Social Moderates ($\beta = -0.257$, $p = 0.021$). This reflects a traditional cultural cleavage where progressive values align with exotic, authentic, and culinary-forward preferences.
* **Economic Political Ideology (Non-Monotonic U-Shape):** In contrast, economic ideology shares a U-shaped relationship with Foodie Taste. Both Economic Conservatives ($\beta = +0.261$, $p = 0.015$) and Economic Liberals ($\beta = +0.316$, $p = 0.0016$) score significantly higher on Foodie Taste than Economic Moderates. This suggests that holding a coherent, defined economic ideology (regardless of whether it is on the left or right pole) is associated with higher cultural foodie capital and distinction.

The code chunk below calculates these predicted scores and displays them side-by-side with 95% confidence intervals.

```{r}
#| label: plot-political-contrast
#| code-summary: "Contrasting effects of social and economic ideology on Dimension 3"
#| fig-width: 10
#| fig-height: 5
#| fig-cap: "Faceted comparison of predicted scores on Dimension 3 (Foodie Taste) by social and economic political ideology, showing a monotonic relationship for social issues and a non-monotonic U-shaped relationship for economic issues."

# Fit Dimension 3 main effects model
fit_d3_main <- lm(Dim.3 ~ race.f + gend.f + sex.f + educ + income + age + spol.f + epol.f, data = mfa.dat)

# Calculate predictions using ggeffects
pred_spol_d3 <- ggpredict(fit_d3_main, terms = "spol.f")
pred_epol_d3 <- ggpredict(fit_d3_main, terms = "epol.f")

# Format data frames
df_spol_d3 <- as.data.frame(pred_spol_d3) |> 
  mutate(
    Ideology_Type = "Social Issues",
    Category = case_when(
      x == "Soc. Con." ~ "Conservative",
      x == "Soc. Mod." ~ "Moderate",
      x == "Soc. Lib." ~ "Liberal"
    )
  )

df_epol_d3 <- as.data.frame(pred_epol_d3) |> 
  mutate(
    Ideology_Type = "Economic Issues",
    Category = case_when(
      x == "Econ. Con." ~ "Conservative",
      x == "Econ. Mod." ~ "Moderate",
      x == "Econ. Lib." ~ "Liberal"
    )
  )

# Combine datasets and order levels
df_plot_d3 <- rbind(df_spol_d3, df_epol_d3) |> 
  mutate(
    Category = factor(Category, levels = c("Liberal", "Moderate", "Conservative")),
    Ideology_Type = factor(Ideology_Type, levels = c("Social Issues", "Economic Issues"))
  )

# Plot side-by-side
p_political_contrast_d3 <- ggplot(df_plot_d3, aes(x = Category, y = predicted, group = Ideology_Type)) +
  geom_line(linewidth = 1, color = "gray50") +
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high, color = Category), size = 0.8, linewidth = 1) +
  facet_wrap(~Ideology_Type, scales = "free_x") +
  scale_color_manual(values = c("Liberal" = "#3498db", "Moderate" = "#2ecc71", "Conservative" = "#e74c3c")) +
  labs(
    title = "Contrasting Effects of Ideology on Foodie Taste (MFA Dim 3)",
    subtitle = "Predicted scores with 95% confidence intervals, adjusted for demographics",
    x = "Ideological Placement",
    y = "Predicted Score (Dimension 3)"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    strip.text = element_text(size = 14, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    axis.text = element_text(size = 12),
    axis.text.x = element_text(size = 11, margin = margin(t = 5)),
    axis.title = element_text(size = 13, face = "bold"),
    panel.spacing = unit(1.5, "lines")
  )

# Return the plot directly to the report
p_political_contrast_d3

# Save plot to Plots directory
ggsave(here::here("Plots", "political_contrast_dim3.png"), p_political_contrast_d3, width = 10, height = 5, dpi = 300)
```

---

# Food subcultures and life course trends (HCPC)

We use Hierarchical Clustering on Principal Components (HCPC) to group individuals into homogenous "taste profiles" or cultural food subcultures based on their MFA coordinates.

```{r}
#| label: run-hcpc
#| code-summary: "Execute HCPC"
hcpc_res <- HCPC(mfa.res, nb.clust = 3, graph = FALSE)
mfa.dat$Cluster <- hcpc_res$data.clust$clust

# Save clustering tree and map
png(here("Plots", "hcpc_tree.png"), width = 800, height = 600)
plot(hcpc_res, choice = "tree")
dev.off()

png(here("Plots", "hcpc_map.png"), width = 800, height = 600)
plot(hcpc_res, choice = "map")
dev.off()
```

The hierarchical algorithm identifies **3 optimal clusters**:
1. **Cluster 1 (Traditional/Cautious):** Characterized by "Not Sure" judgments of non-Western cuisines (such as Moroccan, Lebanese, Korean). This cluster is dominated by social Moderates and white respondents.
2. **Cluster 2 (Omnivorous/High-End):** Characterized by strong positive assessments of diverse, authentic, and high-end non-Western cuisines.
3. **Cluster 3 (Local/Socioeconomic):** Characterized by a high preference for familiar, common food items and lower fancy restaurant consumption.

## Predicting cluster membership: Multinomial logit

To model how demographics, social political ideology, and economic political ideology structure membership in these three taste clusters across the life course, we fit a Multinomial Logit Model predicting cluster membership, with **Cluster 1 (Traditional/Cautious)** as the reference category.

```{r}
#| label: run-multinomial
#| code-summary: "Fit multinomial logit model"
multinom_model <- multinom(
    Cluster ~ spol.f + epol.f + gend.f + educ + age, 
    data = mfa.dat, 
    trace = FALSE
)

# Save multinomial logit table as HTML
tab_model(multinom_model, file = here("Tabs", "multinom_table.html"))
```

To easily detect trends and reduce visual clutter, we compute the predicted probabilities of belonging to each of the three clusters across Age and Social Political Ideology, faceting by cluster.

```{r}
#| label: plot-multinomial
#| code-summary: "Plot faceted predicted probabilities"
pred_res <- ggpredict(multinom_model, terms = c("age [all]", "spol.f"))
pred_df <- as.data.frame(pred_res)

p_multinom <- ggplot(pred_df, aes(x = x, y = predicted, color = group)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~response.level, labeller = labeller(response.level = function(x) paste("Cluster", x))) +
    scale_color_manual(values = c("Soc. Lib." = "blue", "Soc. Mod." = "green", "Soc. Con." = "red")) +
    labs(
        title = "Predicted Probabilities of Taste Cluster Membership",
        subtitle = "Faceted by Taste Profile Cluster",
        x = "Age",
        y = "Predicted Probability",
        color = "Social Ideology"
    ) +
    theme_minimal() +
    theme(text = element_text(size = 12))
p_multinom
ggsave(here("Plots", "pred_prob_faceted.png"), p_multinom, width = 10, height = 6)
```

### Key cluster probability trends

The faceted predicted probability plot reveals stark differences across the life course structured by social beliefs:
* **Cluster 1 (Traditional/Cautious):** The probability of joining this cautious cluster **strongly increases with age** for all groups, but is consistently highest for Social Conservatives and Social Moderates, reaching nearly 70% for older Social Conservatives.
* **Cluster 2 (Omnivorous/High-End):** Highly concentrated among **younger Social Liberals** (probability starts above 60% and declines steeply with age). Younger Social Conservatives have a much lower probability (~25%) of belonging to this high-end omnivore subculture.
* **Cluster 3 (Local/Socioeconomic):** Holds stable across age groups, but shows that Social Moderates and Social Conservatives are consistently more aligned with local/common tastes than Social Liberals across all age bands.
