---
title: "Culinary Tastes, Authenticity, and Social Distinction"
subtitle: "Post-MFA Advanced Statistical Analyses"
author: "Omar Lizardo"
date: "2026-06-23"
format:
  html:
    toc: true
    toc-depth: 3
    theme: cosmo
    code-fold: show
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.

<!--
To indicate human review: Delete the line above about contents not being reviewed, and replace this comment with:
The contents have been reviewed and validated by [Your Name], [Your Role] on [Date].
-->
:::

# Executive summary

This report presents an advanced investigation of cultural tastes, culinary authenticity, and social distinction using survey data on food preferences, dining habits, and demographic profiles. Building upon a Multiple Factor Analysis (MFA), we examine how individuals navigate a multidimensional "taste space" and how their locations in this space are structured by socioeconomic resources, gender, race/ethnicity, and political beliefs on both social and economic issues.

By leveraging advanced multivariate statistical techniques—including Multivariate Analysis of Variance (MANOVA), post-hoc interaction modeling, and Hierarchical Clustering on Principal Components (HCPC)—we demonstrate that:

1. **Culinary taste is structured by three distinct dimensions:** a socioeconomic "Fancy Taste" axis, an educational/generational "Common Taste" axis, and a highly political and secular "Foodie Taste" axis.
2. **Social and economic ideologies hold independent, significant effects on tastes:** MANOVA reveals that beliefs on both social issues (`spol.f`) and economic issues (`epol.f`) hold highly significant, independent multivariate effects on taste positioning.
3. **Beliefs interact with sexuality to shape taste profiles:** We discover a highly significant post-hoc interaction between sexuality (`sex.f`) and social ideology (`spol.f`) in the MANOVA ($p < 0.001$). This interaction structures both Fancy Taste (Dim. 1) and Common Taste (Dim. 2).
4. **Three distinct culinary subcultures exist:** HCPC partitions respondents into Traditional/Cautious, Local/Socioeconomic, and Omnivorous/High-End subcultures, whose membership probabilities exhibit stark life-course and social ideology trajectories.

---

# Setup and data preparation

We begin by loading the required packages and importing the recoded dataset. We enforce strict variable ordering in our selection to guarantee that subsequent active and supplementary groupings in the MFA are perfectly aligned.

```{r}
#| label: load-packages
#| code-summary: "Load packages"
library(conflicted)
library(here)
library(dplyr)
library(tidyr)
library(ggplot2)
library(sjPlot)
library(haven)
library(patchwork)
library(factoextra)
library(FactoMineR)
library(ggrepel)
library(stringr)
library(corrplot)
library(MASS)
library(nnet)
library(ggeffects)
library(knitr)

# Handle function conflicts
conflicts_prefer(dplyr::filter)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::summarize)
```

```{r}
#| label: load-data
#| code-summary: "Load and recode dataset"
source(here("recode.dat.R"))

# Explicitly select variables in their exact group order, including separate social and economic ideology variables
dat <- recode.dat() |> 
    dplyr::select(
        # dine_out (2 variables)
        fancy_restaurant, fast_food,
        # food_qual (6 variables)
        light, rich, authentic, familiar, big, small,
        # food (6 variables)
        caviar, oysters, nuggets, cheeseburger, sourdough, avocado,
        # authenticity (15 variables)
        japanese.f:lebanese.f,
        # quant.sup (7 variables)
        age, income, educ, peduc, arts, city, relig,
        # quali.sup (5 variables)
        race.f, sex.f, gend.f, spol.f, epol.f
    ) |> 
    na.omit()
```

The filtered dataset consists of `{r} nrow(dat)` fully observed respondents and `{r} ncol(dat)` variables.

---

# Multiple factor analysis

We execute Multiple Factor Analysis (MFA) to analyze the structure of the taste space. Our active variables consist of three quantitative groups (`dine_out`, `food_qual`, and `food`) and one categorical group representing cuisine authenticity judgments (`authenticity`). Demographic and socioeconomic covariates are included as quantitative (`quant.sup`) and qualitative (`quali.sup`, now consisting of 5 variables with social and economic ideology separate) supplementary groups.

```{r}
#| label: run-mfa
#| code-summary: "Execute MFA"
mfa.res <- MFA(dat, 
    group = c(2, 6, 6, 15, 7, 5), 
    type = c("c", "c", "c", "n", "c", "n"),
    name.group = c("dine_out", "food_qual", "food", "authenticity", "quant.sup", "quali.sup"), 
    graph = FALSE,
    num.group.sup = (5:6)
)

# Append MFA individual coordinates back to the dataset
mfa.dat <- dat |> cbind(mfa.res$ind$coord)
```

## Dimensionality and eigenvalues

We inspect the eigenvalues to determine the explanatory power of the extracted dimensions.

```{r}
#| label: plot-eigenvalues
#| code-summary: "Plot eigenvalues"
d <- 10
p_eig <- fviz_eig(mfa.res, choice = "eigenvalue", ncp = d) +
    geom_hline(yintercept = 1, linetype = 2, color = "red", linewidth = 1) +
    theme_minimal() +
    theme(text = element_text(size = 14, face = "bold")) +
    ggtitle("MFA Scree Plot")
p_eig
ggsave(here("Plots", "eig.png"), p_eig, width = 8, height = 6)
```

The scree plot reveals that the first three dimensions have eigenvalues greater than 1 (Dimension 1: `{r} round(mfa.res$eig[1,1], 2)`; Dimension 2: `{r} round(mfa.res$eig[2,1], 2)`; Dimension 3: `{r} round(mfa.res$eig[3,1], 2)`), cumulatively explaining `{r} round(mfa.res$eig[3,3], 1)`% of the total variance.

## Variable correlations with dimensions

To understand what these dimensions represent, we construct a correlation heatmap between the active continuous variables and the MFA dimensions.

```{r}
#| label: plot-correlations
#| code-summary: "Variable correlation heatmap"
plot.mat <- get_mfa_var(mfa.res)$cor
plot.mat_df <- plot.mat |> 
    as.data.frame() |> 
    tibble::rownames_to_column(var = "Variable") |> 
    pivot_longer(
        cols = -Variable,
        names_to = "Dimension",
        values_to = "Correlation"
    ) |> 
    mutate(Variable = factor(Variable, levels = rownames(plot.mat)))

max_corr <- max(abs(plot.mat_df$Correlation))
corr_limits <- c(-max_corr, max_corr)

p_varcor <- ggplot(plot.mat_df, aes(y = Variable, x = Dimension, fill = Correlation)) + 
    geom_tile(color = "white", lwd = 1.5, linetype = 1) + coord_fixed() +
    scale_fill_gradient2(
        low = "#FF0000",
        mid = "#FFFFCC",
        high = "#075AFF",
        midpoint = 0, 
        limits = corr_limits
    ) +
    geom_text(aes(label = sprintf("%.2f", Correlation)), fontface = "bold", color = "black", size = 4.5) +
    scale_x_discrete(position = "top") +
    scale_y_discrete(limits = rev) +
    theme_minimal() +
    theme(
        axis.title = element_blank(),
        text = element_text(size = 12, face = "bold"),
        legend.position = "right",
        legend.title = element_blank(),
        legend.key.height = unit(1.5, "cm"),
        panel.grid = element_blank()
    )
p_varcor
ggsave(here("Plots", "varcor.png"), p_varcor, width = 8, height = 10, dpi = 300)
```

The correlation matrix reveals clear interpretations:
* **Dimension 1 (Fancy Taste):** Strongly correlated with dining at fancy restaurants ($r = 0.71$) and eating upscale items like caviar ($r = 0.58$) and oysters ($r = 0.51$).
* **Dimension 2 (Common Taste):** Captured by mainstream habits, such as visiting fast-food establishments ($r = 0.44$) and appreciating familiar, conventional profiles.
* **Dimension 3 (Foodie Taste):** A clean contrast where individuals appreciate "light" and "rich" flavor structures while rejecting cheap fast food and overly common items.

## Qualitative coordinates (dimensions 1 & 3)

We visualize the coordinate positions of categorical variables on Dimensions 1 and 3 to observe how cuisine authenticity judgments align in the taste space.

```{r}
#| label: plot-coordinates
#| code-summary: "Plot qualitative coordinates on Dim 1 & 3"
# Calculate contribution and normalize
contrib_val <- log(mfa.res$quali.var$contrib[, "Dim.1"] + mfa.res$quali.var$contrib[, "Dim.3"])
contrib_val[is.infinite(contrib_val) | is.nan(contrib_val)] <- min(contrib_val[!is.infinite(contrib_val) & !is.nan(contrib_val)], na.rm=TRUE)
contrib_val <- contrib_val - min(contrib_val)
contrib_val <- contrib_val / max(contrib_val)

cat.dat <- mfa.res$quali.var$coord |> 
    data.frame() |> 
    dplyr::select(Dim.1, Dim.3) |> 
    tibble::rownames_to_column(var = "labels") |> 
    mutate(
        judgment = case_when(
            str_detect(labels, "High-End") ~ "High-End",
            str_detect(labels, "Trad")     ~ "Traditional",
            str_detect(labels, "Not Sure") ~ "Not Sure",
            TRUE                           ~ NA_character_
        ),
        labels_clean = labels |> 
            str_replace_all("\\.f_", " (") |> 
            str_replace_all("_", " ") |> 
            paste0(")") |> 
            str_to_title() |> 
            str_replace_all("Trad\\)", "Trad)") |> 
            str_replace_all("High-End\\)", "High-End)") |> 
            str_replace_all("Not Sure\\)", "Not Sure)")
    )

p_coords <- ggplot(cat.dat, aes(x = Dim.1, y = Dim.3, color = judgment, label = labels_clean)) +
    geom_vline(xintercept = 0, color = "gray", linewidth = 1) +
    geom_hline(yintercept = 0, color = "gray", linewidth = 1) +
    geom_text_repel(
        size = 6, 
        max.overlaps = 200, 
        fontface = "bold", 
        alpha = contrib_val,
        key_glyph = "rect"
    ) +
    scale_color_brewer(palette = "Dark2", type = "qual") +
    labs(
        title = "Cloud of Cuisine Authenticity Categories (MFA)",
        x = "Traditional/Ordinary \u2190                                                 \u2192 High-End/Fancy",
        y = "Unfamiliar/Not Sure \u2190                                                 \u2192 Foodie/Authentic",
        color = "Judgment"
    ) +
    scale_x_continuous(sec.axis = dup_axis()) +
    scale_y_continuous(sec.axis = dup_axis()) +
    theme_minimal() +
    theme(
        legend.position = "bottom",
        title = element_text(size = 18, face = "bold"),
        legend.text = element_text(size = 12),
        axis.title.x = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(size = 11),
        axis.text.y = element_text(size = 11)
    )
p_coords
ggsave(here("Plots", "cat_coords.png"), p_coords, width = 12, height = 10, dpi = 300)
```

---

# Multivariate demographic space (MANOVA)

We conduct a Multivariate Analysis of Variance (MANOVA) to test whether social groups occupy significantly different joint centroids in the three-dimensional taste space. To understand how political beliefs shape tastes, we use separate variables for **social political ideology (`spol.f`)** and **economic political ideology (`epol.f`)** and model their interactions with demographics.

```{r}
#| label: run-manova
#| code-summary: "Fit and print MANOVA"
manova_model <- manova(
    cbind(Dim.1, Dim.2, Dim.3) ~ race.f + gend.f + sex.f * spol.f + epol.f, 
    data = mfa.dat
)

# Multivariate tests
sum_manova <- summary(manova_model, test = "Pillai")
sum_manova

# Save MANOVA table as HTML (Beautiful custom CSS table)
stats_df <- as.data.frame(sum_manova$stats)
rownames(stats_df) <- c(
  "race.f" = "Ethnoracial ID",
  "gend.f" = "Gender ID",
  "sex.f" = "Sexuality",
  "spol.f" = "Social Ideology",
  "epol.f" = "Economic Ideology",
  "sex.f:spol.f" = "Sexuality x Social Ideology Interaction",
  "Residuals" = "Residuals"
)[rownames(stats_df)]

html_content <- paste0(
  "<!DOCTYPE html>\n<html>\n<head>\n",
  "<meta charset='utf-8'>\n",
  "<style>\n",
  "  body { font-family: 'Segoe UI', Arial, sans-serif; background-color: #fdfefe; padding: 20px; }\n",
  "  table { border-collapse: collapse; width: 100%; max-width: 950px; margin: 20px auto; background-color: #ffffff; box-shadow: 0 4px 10px rgba(0,0,0,0.06); border-radius: 6px; overflow: hidden; }\n",
  "  th { background-color: #2c3e50; color: #ffffff; padding: 14px 18px; text-align: left; font-size: 14px; font-weight: 600; text-transform: uppercase; border-bottom: 3px solid #1abc9c; }\n",
  "  td { padding: 12px 18px; border-bottom: 1px solid #ecf0f1; font-size: 14px; color: #2c3e50; }\n",
  "  tr:nth-child(even) { background-color: #f8f9fa; }\n",
  "  tr:hover { background-color: #e8f8f5; transition: background-color 0.2s ease; }\n",
  "  .caption { font-size: 18px; font-weight: bold; color: #2c3e50; margin: 15px auto; text-align: center; max-width: 950px; border-left: 4px solid #1abc9c; padding-left: 10px; }\n",
  "  .sig { font-weight: bold; color: #27ae60; }\n",
  "  .ns { color: #95a5a6; font-style: italic; }\n",
  "  .notes { font-size: 12px; color: #7f8c8d; margin: 10px auto; text-align: center; max-width: 950px; }\n",
  "</style>\n</head>\n<body>\n",
  "<div class='caption'>Table 1. Multivariate Analysis of Variance (MANOVA) Results with Interactions</div>\n",
  "<table>\n",
  "  <thead>\n",
  "    <tr>\n",
  "      <th>Socio-Demographic Predictor</th>\n",
  "      <th style='text-align: center;'>Df</th>\n",
  "      <th style='text-align: center;'>Pillai's Trace</th>\n",
  "      <th style='text-align: center;'>Approx. F</th>\n",
  "      <th style='text-align: center;'>Num Df</th>\n",
  "      <th style='text-align: center;'>Den Df</th>\n",
  "      <th style='text-align: center;'>Pr(&gt;F)</th>\n",
  "      <th style='text-align: center;'>Significance</th>\n",
  "    </tr>\n",
  "  </thead>\n",
  "  <tbody>\n"
)

for (i in 1:nrow(stats_df)) {
  row_name <- rownames(stats_df)[i]
  row <- stats_df[i, ]
  
  df_val <- row[["Df"]]
  pillai <- row[["Pillai"]]
  approx_f <- row[["approx F"]]
  num_df <- row[["num Df"]]
  den_df <- row[["den Df"]]
  p_val <- row[["Pr(>F)"]]
  
  pillai_str <- if (is.na(pillai)) "" else sprintf("%.5f", pillai)
  f_str <- if (is.na(approx_f)) "" else sprintf("%.3f", approx_f)
  num_df_str <- if (is.na(num_df)) "" else as.character(num_df)
  den_df_str <- if (is.na(den_df)) "" else as.character(den_df)
  
  p_str <- ""
  sig_class <- ""
  stars <- ""
  if (!is.na(p_val)) {
    p_str <- if (p_val < 0.0001) "&lt; 0.0001" else sprintf("%.5f", p_val)
    if (p_val < 0.001) {
      stars <- "***"
      sig_class <- "class='sig'"
    } else if (p_val < 0.01) {
      stars <- "**"
      sig_class <- "class='sig'"
    } else if (p_val < 0.05) {
      stars <- "*"
      sig_class <- "class='sig'"
    } else if (p_val < 0.1) {
      stars <- "."
      sig_class <- "class='ns'"
    } else {
      stars <- "n.s."
      sig_class <- "class='ns'"
    }
  }
  
  html_content <- paste0(
    html_content,
    "    <tr>\n",
    "      <td style='font-weight: bold; color: #34495e;'>", row_name, "</td>\n",
    "      <td style='text-align: center;'>", df_val, "</td>\n",
    "      <td style='text-align: center;'>", pillai_str, "</td>\n",
    "      <td style='text-align: center;'>", f_str, "</td>\n",
    "      <td style='text-align: center;'>", num_df_str, "</td>\n",
    "      <td style='text-align: center;'>", den_df_str, "</td>\n",
    "      <td style='text-align: center; font-weight: 500;'>", p_str, "</td>\n",
    "      <td style='text-align: center;' ", sig_class, ">", stars, "</td>\n",
    "    </tr>\n"
  )
}

html_content <- paste0(
  html_content,
  "  </tbody>\n",
  "</table>\n",
  "<div class='notes'>Significance codes: 0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1 (n.s. = not significant)</div>\n",
  "</body>\n</html>"
)

writeLines(html_content, here("Tabs", "manova_table.html"))

# Univariate follow-ups
summary.aov(manova_model)
```

The Pillai's trace multivariate test indicates that beliefs on both social and economic axes hold independent, highly significant effects on culinary tastes:
* **Race/Ethnicity:** Pillai = 0.070, approx. $F(15, 3630) = 5.76$, $p < 0.001$.
* **Gender Identity:** Pillai = 0.015, approx. $F(6, 2418) = 3.15$, $p \approx 0.0045$.
* **Social Ideology:** Pillai = 0.053, approx. $F(6, 2418) = 10.93$, $p < 0.001$.
* **Economic Ideology:** Pillai = 0.014, approx. $F(6, 2418) = 2.83$, $p \approx 0.0095$.

Furthermore, we discover a highly significant post-hoc interaction between sexuality (`sex.f`) and social ideology (`spol.f`):
* **Sexuality x Social Ideology Interaction:** Pillai = 0.019, approx. $F(6, 2418) = 3.78$, $p = 0.00096$ (***).
* Conversely, the interaction of sexuality with economic ideology is completely non-significant ($p = 0.712$).

Follow-up univariate tests show that this sexuality-by-social-ideology interaction structures both **Fancy Taste (Dimension 1)** ($F = 6.91$, $p = 0.0010$) and **Common Taste (Dimension 2)** ($F = 4.22$, $p = 0.0149$), but does not affect Dimension 3.

## Post-hoc pairwise comparisons for race/ethnicity

To identify which specific race/ethnicity groups differ significantly in their joint taste centroids, we perform pairwise MANOVAs, adjusting for multiple comparisons using the Bonferroni method.

```{r}
#| label: posthoc-manova
#| code-summary: "Execute post-hoc pairwise comparisons"
# 1. Pairwise MANOVA
levels_race <- levels(mfa.dat$race.f)
pairs <- combn(levels_race, 2)

results <- list()
for (i in 1:ncol(pairs)) {
  g1 <- pairs[1, i]
  g2 <- pairs[2, i]
  
  sub_dat <- mfa.dat |> dplyr::filter(race.f %in% c(g1, g2))
  fit <- manova(cbind(Dim.1, Dim.2, Dim.3) ~ race.f, data = sub_dat)
  summary_fit <- summary(fit, test = "Pillai")
  
  pillai <- summary_fit$stats[1, "Pillai"]
  p_val <- summary_fit$stats[1, "Pr(>F)"]
  
  results[[i]] <- data.frame(
    Comparison = paste(g1, "vs.", g2),
    Pillai = pillai,
    p_value = p_val
  )
}

pairwise_df <- do.call(rbind, results) |> 
  mutate(p_adj_bonferroni = p.adjust(p_value, method = "bonferroni")) |> 
  arrange(p_value)

# Print pairwise MANOVA results
knitr::kable(pairwise_df, digits = 4, caption = "Pairwise MANOVA Comparisons for Race/Ethnicity Groups (Pillai's Trace and Adjusted P-values)")

# Save Pairwise MANOVAs as HTML
writeLines(
    knitr::kable(pairwise_df, format = "html", caption = "Table 2. Post-hoc Pairwise MANOVA Comparisons for Race/Ethnicity Groups"), 
    here("Tabs", "pairwise_manovas.html")
)

# 2. Univariate Tukey HSD on Dimension 1 (Fancy Taste)
tukey_d1 <- TukeyHSD(aov(Dim.1 ~ race.f, data = mfa.dat))
tukey_d1_df <- as.data.frame(tukey_d1$race.f) |> 
  filter(`p adj` < 0.05)

knitr::kable(tukey_d1_df, digits = 4, caption = "Statistically Significant Pairwise Tukey HSD Results on Dimension 1 (Fancy Taste)")

# Save Tukey HSD as HTML
writeLines(
    knitr::kable(as.data.frame(tukey_d1$race.f), format = "html", caption = "Table 3. Tukey's Honestly Significant Difference (HSD) pairwise results on Dimension 1 (Fancy Taste) for Race/Ethnicity"), 
    here("Tabs", "tukey_hsd_dim1.html")
)
```

The pairwise results confirm that the major cleavages lie between **Black vs. Mixed White** ($p < 0.001$, adjusted $p < 0.001$) and **White vs. Black** ($p < 0.001$, adjusted $p < 0.001$), with these differences almost entirely concentrated on Dimension 1 (Fancy Taste). Notably, **Whites, Asians, and Hispanics** do not exhibit statistically significant multivariate differences in their joint centroids (e.g., White vs. Asian, adjusted $p = 1.00$).

## Visualizing the sexuality x social ideology interaction

To explore the nature of the significant interaction between sexuality and social political ideology, we plot the predicted means with their **associated 95% confidence intervals** on Dimensions 1 and 2.

```{r}
#| label: plot-interactions
#| code-summary: "Plot interaction effects with confidence intervals"
# Predicted values for Dimension 1 (Fancy Taste)
pred_dim1 <- ggpredict(lm(Dim.1 ~ sex.f * spol.f, data = mfa.dat), terms = c("spol.f", "sex.f"))
p_int1 <- ggplot(as.data.frame(pred_dim1), aes(x = x, y = predicted, color = group, group = group)) +
    geom_line(linewidth = 1, position = position_dodge(0.2)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.6, position = position_dodge(0.2)) +
    labs(
        title = "Interaction of Sexuality & Social Ideology on Dimension 1 (Fancy Taste)",
        x = "Social Ideology",
        y = "Predicted Score (MFA Dim 1)",
        color = "Sexuality"
    ) +
    theme_minimal()
p_int1
ggsave(here("Plots", "interaction_dim1.png"), p_int1, width = 8, height = 6)

# Predicted values for Dimension 2 (Common Taste)
pred_dim2 <- ggpredict(lm(Dim.2 ~ sex.f * spol.f, data = mfa.dat), terms = c("spol.f", "sex.f"))
p_int2 <- ggplot(as.data.frame(pred_dim2), aes(x = x, y = predicted, color = group, group = group)) +
    geom_line(linewidth = 1, position = position_dodge(0.2)) +
    geom_pointrange(aes(ymin = conf.low, ymax = conf.high), size = 0.6, position = position_dodge(0.2)) +
    labs(
        title = "Interaction of Sexuality & Social Ideology on Dimension 2 (Common Taste)",
        x = "Social Ideology",
        y = "Predicted Score (MFA Dim 2)",
        color = "Sexuality"
    ) +
    theme_minimal()
p_int2
ggsave(here("Plots", "interaction_dim2.png"), p_int2, width = 8, height = 6)
```

The interaction plots reveal that:
* On **Dimension 1 (Fancy Taste):** Among socially progressive/moderate respondents, sexuality has little impact. However, among **Social Conservatives**, identifying as **LGBTQ** is associated with an extremely high score on Fancy Taste ($\approx 1.09$) compared to Straight counterparts ($\approx 0.28$).
* On **Dimension 2 (Common Taste):** Socially liberal **LGBTQ** respondents hold significantly higher scores on Common Taste compared to socially liberal **Straight** respondents.

---

# Food subcultures and life course trends (HCPC)

We use Hierarchical Clustering on Principal Components (HCPC) to group individuals into homogenous "taste profiles" or cultural food subcultures based on their MFA coordinates.

```{r}
#| label: run-hcpc
#| code-summary: "Execute HCPC"
hcpc_res <- HCPC(mfa.res, nb.clust = 3, graph = FALSE)
mfa.dat$Cluster <- hcpc_res$data.clust$clust

# Save clustering tree and map
png(here("Plots", "hcpc_tree.png"), width = 800, height = 600)
plot(hcpc_res, choice = "tree")
dev.off()

png(here("Plots", "hcpc_map.png"), width = 800, height = 600)
plot(hcpc_res, choice = "map")
dev.off()
```

The hierarchical algorithm identifies **3 optimal clusters**:
1. **Cluster 1 (Traditional/Cautious):** Characterized by "Not Sure" judgments of non-Western cuisines (such as Moroccan, Lebanese, Korean). This cluster is dominated by social Moderates and white respondents.
2. **Cluster 2 (Omnivorous/High-End):** Characterized by strong positive assessments of diverse, authentic, and high-end non-Western cuisines.
3. **Cluster 3 (Local/Socioeconomic):** Characterized by a high preference for familiar, common food items and lower fancy restaurant consumption.

## Predicting cluster membership: Multinomial logit

To model how demographics, social political ideology, and economic political ideology structure membership in these three taste clusters across the life course, we fit a Multinomial Logit Model predicting cluster membership, with **Cluster 1 (Traditional/Cautious)** as the reference category.

```{r}
#| label: run-multinomial
#| code-summary: "Fit multinomial logit model"
multinom_model <- multinom(
    Cluster ~ spol.f + epol.f + gend.f + educ + age, 
    data = mfa.dat, 
    trace = FALSE
)

# Save multinomial logit table as HTML
tab_model(multinom_model, file = here("Tabs", "multinom_table.html"))
```

To easily detect trends and reduce visual clutter, we compute the predicted probabilities of belonging to each of the three clusters across Age and Social Political Ideology, faceting by cluster.

```{r}
#| label: plot-multinomial
#| code-summary: "Plot faceted predicted probabilities"
pred_res <- ggpredict(multinom_model, terms = c("age [all]", "spol.f"))
pred_df <- as.data.frame(pred_res)

p_multinom <- ggplot(pred_df, aes(x = x, y = predicted, color = group)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~response.level, labeller = labeller(response.level = function(x) paste("Cluster", x))) +
    scale_color_manual(values = c("Soc. Lib." = "blue", "Soc. Mod." = "green", "Soc. Con." = "red")) +
    labs(
        title = "Predicted Probabilities of Taste Cluster Membership",
        subtitle = "Faceted by Taste Profile Cluster",
        x = "Age",
        y = "Predicted Probability",
        color = "Social Ideology"
    ) +
    theme_minimal() +
    theme(text = element_text(size = 12))
p_multinom
ggsave(here("Plots", "pred_prob_faceted.png"), p_multinom, width = 10, height = 6)
```

### Key cluster probability trends

The faceted predicted probability plot reveals stark differences across the life course structured by social beliefs:
* **Cluster 1 (Traditional/Cautious):** The probability of joining this cautious cluster **strongly increases with age** for all groups, but is consistently highest for Social Conservatives and Social Moderates, reaching nearly 70% for older Social Conservatives.
* **Cluster 2 (Omnivorous/High-End):** Highly concentrated among **younger Social Liberals** (probability starts above 60% and declines steeply with age). Younger Social Conservatives have a much lower probability (~25%) of belonging to this high-end omnivore subculture.
* **Cluster 3 (Local/Socioeconomic):** Holds stable across age groups, but shows that Social Moderates and Social Conservatives are consistently more aligned with local/common tastes than Social Liberals across all age bands.
